A Survey of Scientific Workflow Management Systems

### 2014-05-19 paper.tex
bib fixes
Taverna text

    \item \textbf{Interoperability.} Interoperable workflows will benefit to a
        broad audience of users. Many efforts at this are made in the recent
        past, such as by the consurtium projects and by individual players.
        However, a truly interoperable set of standards for workflows are yet
        to emerge.
\end{enumerate}

\section{Objectives and Organization}
% What are the goals of this work?
% How is this paper supposed to help:
% -- Existing scientific users who have not adopted workflows
% -- Existing users who have adopted workflows
% -- New users
% -- Experts in the field

@@

 \item \textbf{Catalog.} We study a number of important features as offered by
     contemporary SWEs. This will help catalog and collect the SWEs by their
     offerings and identify mechanisms and needs for new and existing features.
     This will also help identify current trends in scientific computing
     community.

@@

A Taverna workflow consists of a collection of processors connected by data
links, which establish a dependency between the output(s) of a processor and
the input(s) of another. Apart from data links, the processors may also be
connected through coordination links that make a processor invoke when a
previous processor ends its execution.  The processors are of different kinds
depending on the application code to be invoked: typically being services, with
input and output \textit{ports} that correspond to the operation's invocations
and response sequences defined in the service`s interface expressed as WSDL.
Java classes and local scripts can be used as workflow processors, as long as
they expose their signature as input and output ports as required by the model.
The engine orchestrates the execution of the processors in a way that is
consistent with the dependencies, and manages the flow of data through the
processors' ports. The overall workflow execution is data-driven, with a \emph{push}
model: a processor's execution is started as soon as all of its inputs are
available.

Taverna is a flexible and extendible workflow environment
\cite{missier-soiland-reyes-etal:2010,sroka2009a} in that extending the
engine's functionality is facilitated through a Service Provider Interface
(SPI). A third party developer intending to extend a functionality must write
its logic as an SPI and publish it as an extension SPI. This may manifest
itself in different ways depending upon the intended requirements of the
developer, examples being as a task in the service panel or internally within
the enactor as additional functionality.

The Taverna engine implements a sophisticated mechanism for enactment of
parallel workflows. These mechanisms manifest in the form of pipelines and an
advanced superpipelined enactment of workflows. In the case of pipelined
execution, a simple pipeline of tasks are overlapped into a back to back
execution for a single data set. Such a mechanism induces multiple downstream
workflow processors running in parallel for a dataset. In the case of
superpipelines, multiple such pipelines instances run for multi-datasets
providing a high degree of parallelism limited only by the availability of
underlying hardware resources.

Taverna is not natively interfaced with external computational infrastructure but the same can
be achieved with the help of extending the engine through one of its SPIs~\cite{maheshwari-missier-etal:2009}. 
\km{ToDo: Taverna current activity, applications support, GPU, Cloud support}

@@

%
\paragraph{Swift/Script, Tcl} Swift \cite{swift} is a script-based scientific
workflow environment developed at the University of Chicago and Argonne
National Laboratory. A Swift workflow specified in the Swift scripting language
and is translated to an intermediate XML-based representation and enacted by
the Karajan workflow enactment engine. The SwiftScript language is proposed as
a workflow language founded on a scripting approach that programmers are
familiar with. To make the language data parallel, Swift script generalizes the
use of futures \cite{Halstead85multilisp:a} for every variables defined in the
script.  Futures are non-blocking assignment variables, using a proxy to ensure
immediate execution of assignment and performing lazy blocking on variable
value accesses only. The systematic use of futures for every variables in the
language makes it completely data-driven: the execution progresses
asynchronously as long as a data access is not blocking. The availability of
data enables blocked thread to restart execution. Swift facilitates a highly
asynchronous workflow execution. Swift supports advanced workflow constructs
such as parallel loops and conditionals. Furthermore Swift supports processing
of explicitly defined arrays of data. Swift workflows take advantage of
POSIX-style shared file system to reuse the data produced by other processors.
Recent advances in Swift:
%Recent advances in Swift: 
\begin{enumerate}
    \item Swift has been shown to run on multiple
        clouds~\cite{ccgrid2013} via its coasters mechanism~\cite{ucc2011}.

>>>>

    \item \textbf{Interoperability.} Interoperable workflows will benefit to a
        broad audience of users. Many efforts at this are made in the recent
        past, such as by the consurtium projects and by individual players.
        However, a truly interoperable set of standards for workflows are yet
        to emerge.
\end{enumerate}
\section{Objectives and Organization}
% What are the goals of this work?
% How is this paper supposed to help:
% -- Existing scientific users who have not adopted workflows
% -- Existing users who have adopted workflows
% -- New users
% -- Experts in the field

@@

 \item \textbf{Catalog.} We study a number of important features as offered by
     contemporary SWEs. This will help catalog and collect the SWEs by their
     offerings and identify mechanisms and needs for new and existing features
     in one place.  This will also help identify current trends and approaches
     to computation in scientific communities.

@@

A Taverna workflow consists of a collection of processors connected by data and
coordination links, which establish a dependency between the output(s) of a
processor and the input(s) of another. The processors are of different kinds
depending on the application code to be invoked: typically being services, with
input and output \textit{ports} that correspond to the operation's invocations
and response sequences defined in the service`s interface expressed as REST or WSDL.
Java classes and local scripts can be used as workflow processors, as long as
they expose their signature as input and output ports as required by the model.
The engine orchestrates the execution of the processors in a way that is
consistent with the dependencies, and manages the flow of data through the
processors' ports. The overall workflow execution is data-driven, with a
\emph{push} model: a processor's execution is started as soon as all of its
inputs are available.

Taverna is a flexible and extendible workflow environment
\cite{missier-soiland-reyes-etal:2010,sroka2009a} in that extending the
engine's functionality is facilitated through a Service Provider Interface
(SPI). The Taverna engine implements a sophisticated mechanism for enactment of
parallel workflows. These mechanisms manifest in the form of complex pipelines
and an advanced superpipelined enactment of workflows. Such a mechanism induces
multiple downstream workflow processors running in parallel for a dataset. In
the case of superpipelines, multiple pipeline instances run for multi-datasets
providing a high degree of parallelism limited only by the availability of
underlying resources. Taverna is not natively interfaced with external
computational infrastructure but the same can be achieved with the help of
extending the engine through one of its
SPIs~\cite{maheshwari-missier-etal:2009}. Recent implementations of Taverna
offers a client-server based environment in which a Taverna server can remotely
serve the workflow composition and execution services. This mode is useful in
the cloud computing environments where the server decides where to run the
Taverna services in the cloud.

@@

%
\paragraph{Swift/Script, Tcl} Swift \cite{swift} is a script-based scientific
workflow environment developed at the University of Chicago and Argonne
National Laboratory. A Swift workflow specified in the Swift scripting language
and is translated to an intermediate XML-based representation and enacted by
the Karajan workflow enactment engine. The SwiftScript language is proposed as
a workflow language founded on a scripting approach that programmers are
familiar with. To make the language data parallel, Swift script generalizes the
use of futures~\cite{futures} for every variables defined in the
script.  Futures are non-blocking assignment variables, using a proxy to ensure
immediate execution of assignment and performing lazy blocking on variable
value accesses only. The systematic use of futures for every variables in the
language makes it completely data-driven: the execution progresses
asynchronously as long as a data access is not blocking. The availability of
data enables blocked thread to restart execution. Swift facilitates a highly
asynchronous workflow execution. Swift supports advanced workflow constructs
such as parallel loops and conditionals. Furthermore Swift supports processing
of explicitly defined arrays of data. Swift workflows take advantage of
POSIX-style shared file system to reuse the data produced by other processors.
Recent advances in Swift:
%Recent advances in Swift: 
\begin{enumerate}
    \item Swift has been shown to run on multiple
        clouds~\cite{ccgrid2013} via its coasters mechanism~\cite{ucc2011}.
